---
title: 'Lab 3: 12 Days of Christmas'
author: "Stephanie Liu"
date: "10/10/2021"
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: vignette
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(glue)
library(english)
```


## Introduction

**The song “12 Days of Christmas”, written around 1780, tells the tale of many gifts a person receives in the days leading up to Christmas.**

(https://en.wikipedia.org/wiki/The_Twelve_Days_of_Christmas_(song))

**These gifts repeat and compound; on the first day, the narrator recieves**

A partridge in a pear tree.

**On the last day, he or she receives**

Twelve Drummers Drumming

Eleven Pipers Piping

Ten Lords a Leaping

Nine Ladies Waiting

Eight Maids a Milking

Seven Swans a Swimming

Six Geese a Laying

Five Golden Rings

Four Calling Birds

Three French Hens

Two Turtle Doves

And a Partridge in a Pear Tree

**Your task will be to write functions and loop over them to sing the whole song.**

## A Test Dataset

**Run the code below to load a dataset called xmas that contains the crucial information about the gifts in the song. We will use this dataset to test out our functions as we work on them.**

**Remember, though, that future datasets might not have the exact same information as xmas! You’ll want to write your functions to be general; that is, they could be used with any dataset.**

```{r}
xmas <- read.csv("https://www.dropbox.com/s/12ipnczn0fygglt/xmas.csv?dl=1")
```

**Note that your functions can - and should! - reference each other.**
**That is, don’t duplicate code; use earlier, smaller functions inside your larger functions.**

### Build from small pieces

**If you have some trouble getting started, I recommend writing a function that works in one case, and then trying to generalize.**

**For example, in building my sing_verse function, I might first write a version called sing_third_verse that sings**

On the third day of Christmas, my true love gave to me:

three french hens,

two turtle doves,

and a patridge in a pear tree.

**Then, I would figure out how to let the user input the number and have the function give back the right result.**

### Don’t sweat the small stuff

**There’s a lot you can do to polish up the way the song prints.**

**However, the goal of this lab is to practice writing functions and using iteration. Don’t get bogged down in details like how the song displays, or small grammar rules, until you’ve finished the main tasks.**

### Test every step

**Whenever you write a new line of code, make sure that what it outputs is what you intended! You can use the xmas dataset to test your work.**

## Step One: Make gift phrases

**Write a function called make_phrase that takes as input the necessary information, and returns a phrase. For example,**


```{r, eval=FALSE}
make_phrase(num = 10, 
            gift = "lords a-leaping")
```

should return

"ten lords a-leaping"

*Hint: The function english from the package also called “english” will convert numbers like “10” to words like “ten”*

```{r}
make_phrase <- function(num, gift) {
  if(num == 1){
    word_day <- "a"
  }
  else{
    word_day <- english(num)
  }
  
  phrase <- glue("{word_day} {gift}")
  return(phrase)
  
  }
  
```


```{r}
make_phrase(num = 10, 
            gift = "lords a-leaping")
```

**Try your function out on the xmas data, by making a new variable containing the daily phrases.**

```{r}

xmas %>%
  mutate(
    Gift_Phrase = map2_chr(Day, Gift.Item, ~make_phrase(.x, .y))
  )

```


## Step Two: Verses

**Write a function called sing_verse that takes as input:**

A number indicating which day to sing about

A vector that contains the day labels, like “first”, “second” etc.

A vector that contains the gift for each day.

**For example,**

```{r,eval=FALSE}
sing_verse(2, xmas$Day.in.Words, xmas$Gift.Item)
```

should return

On the second day of Christmas, my true love sent to me,
two turtle doves 
and one partridge in a pear tree.

*Hint: Use your make_phrase function inside sing_verse.*

*Hint 2: Although there exists ways to accomplish this without loops/maps, you may want to try using an iterative approach.*

```{r}
sing_verse <- function(num, day_label, gift) {
  
  verse <- glue("On the {day_label[num]} day of Christmas, my true love gave to me,")
  
  if (num == 1){
    verse <- glue("{verse}
                  {make_phrase(num, gift[num])}")
  }
  else{
    for (i in 1:num) {
     
     if(((num + 1) - i) == 1){
       verse <- glue("{verse} 
                       and {make_phrase((num + 1)-i, gift[(num + 1)-i])}" )
     }
     else{
       verse <- glue("{verse} 
                       {make_phrase((num + 1)-i, gift[(num + 1)-i])}" )
     }
    
  }
  
  }
 
  full_verse <- glue("{verse}")
  return(full_verse)
}
```


```{r}
# test with 6
sing_verse(6, xmas$Day.in.Words, xmas$Gift.Item)
```

```{r}
sing_verse(2, xmas$Day.in.Words, xmas$Gift.Item)
```


## Step Three: Sing the whole song.

**Write a function called sing_xmas_song() that takes as input:**

A vector of day numbers (1, 2, 3, ...)

A vector of day names (first, second, …)

A vector of gifts

And returns the whole text of the song, for all 12 verses.

**This step will require you to use iteration in some way.**

```{r}
sing_xmas_song <- function(nums, day_names, gifts) {
  
  song <- glue("")
  for (i in 1:length(nums)) {
    song <- glue("{song} \n
                 {sing_verse(i, day_names, gifts)}")
    
  }
  return(song)
}

```

```{r}
sing_xmas_song(xmas$Day, xmas$Day.in.Words, xmas$Gift.Item)
```


## Step Four: Try your functions

**Use the following code to load a dataset that is different from the original xmas song. Note that the column names of this dataset are not the same as they were for the original xmas data!**

```{r}
xmas2 <- read.csv("https://www.dropbox.com/s/tmilnd6ad8k26wy/xmas2.csv?dl=1")
```

**Use the function sing_xmas_song() with this dataset to sing a new version of “12 Days of Christmas”.**

```{r}
sing_xmas_song(xmas2$Day, xmas2$Day.in.Words, xmas2$Gift.Item)
```



